---
title: "islr_chapter_5"
author: "Mitch Borgert"
date: "March 31, 2018"
output: 
  html_document: 
    keep_md: yes  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#3. We now review k-fold cross-validation.
##(a) Explain how k-fold cross-validation is implemented.

The k-fold cross validation is implemented by taking the n observations and randomly splitting it into k non-overlapping groups of length of approximately n/k. These groups acts as a validation set, and the remainder (of length n−n/k) acts as a training set. The test error is then estimated by averaging the k resulting MSE estimates.

##(b) What are the advantages and disadvantages of k-fold crossvalidation relative to:
###i. The validation set approach?

The validation estimate of the test error rate can be highly variable depending which observations are included in the training set and which observations are included in the validation set. Also, only a subset of the observations are used to fit the model. Since statistical methods tend to perform worse when trained on fewer observations, this suggests that the validation set error rate may tend to overestimate the test error rate for the model fit on the entire data set.

###ii. LOOCV?

It requires fitting the potentially computationally expensive model n times compared to k-fold cross-validation which requires the model to be fitted only k times. Second, the LOOCV cross-validation approach may give approximately unbiased estimates of the test error, since each training set contains n−1 observations; however, this approach has higher variance than k-fold cross-validation since we are averaging the outputs of n fitted models trained on an almost identical set of observations, these outputs are highly correlated, and the mean of highly correlated quantities has higher variance than less correlated ones. So, there is a bias-variance trade-off associated with the choice of k in k-fold cross-validation.


#5
##(a) Fit a logistic regression model that uses income and balance to predict default.
```{r}
library(ISLR)
attach(Default)
set.seed(1)
fit.glm <- glm(default ~ income + balance, data = Default, family = "binomial")
summary(fit.glm)
```
##(b) Using the validation set approach, estimate the test error of this
model. In order to do this, you must perform the following steps:

###i. Split the sample set into a training set and a validation set.
###ii. Fit a multiple logistic regression model using only the training observations.
```{r}
train <- sample(dim(Default)[1], dim(Default)[1] / 2)
fit.glm <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train)
summary(fit.glm)
```

###iii. Obtain a prediction of default status for each individual in the validation set by computing the posterior probability of default for that individual, and classifying the individual to the default category if the posterior probability is greater than 0.5.

###iv. Compute the validation set error, which is the fraction of the observations in the validation set that are misclassified.
```{r}
probs <- predict(fit.glm, newdata = Default[-train, ], type = "response")
pred.glm <- rep("No", length(probs))
pred.glm[probs > 0.5] <- "Yes"
mean(pred.glm != Default[-train, ]$default)

```

##(c) Repeat the process in (b) three times, using three different splits of the observations into a training set and a validation set. Comment on the results obtained.
```{r}
train <- sample(dim(Default)[1], dim(Default)[1] / 2)
fit.glm <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train)
probs <- predict(fit.glm, newdata = Default[-train, ], type = "response")
pred.glm <- rep("No", length(probs))
pred.glm[probs > 0.5] <- "Yes"
mean(pred.glm != Default[-train, ]$default)

train <- sample(dim(Default)[1], dim(Default)[1] / 2)
fit.glm <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train)
probs <- predict(fit.glm, newdata = Default[-train, ], type = "response")
pred.glm <- rep("No", length(probs))
pred.glm[probs > 0.5] <- "Yes"
mean(pred.glm != Default[-train, ]$default)

train <- sample(dim(Default)[1], dim(Default)[1] / 2)
fit.glm <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train)
probs <- predict(fit.glm, newdata = Default[-train, ], type = "response")
pred.glm <- rep("No", length(probs))
pred.glm[probs > 0.5] <- "Yes"
mean(pred.glm != Default[-train, ]$default)
```

##(d) Now consider a logistic regression model that predicts the probability of default using income, balance, and a dummy variable for student. Estimate the test error for this model using the validation set approach. Comment on whether or not including a dummy variable for student leads to a reduction in the test error rate.

```{r}
train <- sample(dim(Default)[1], dim(Default)[1] / 2)
fit.glm <- glm(default ~ income + balance + student, data = Default, family = "binomial", subset = train)
pred.glm <- rep("No", length(probs))
probs <- predict(fit.glm, newdata = Default[-train, ], type = "response")
pred.glm[probs > 0.5] <- "Yes"
mean(pred.glm != Default[-train, ]$default)
```

